---
title: "gee_upload"
format: html
editor_options: 
  chunk_output_type: console
---

## Quarto

* [Python Installation | Google Earth Engine | Google Developers](https://developers.google.com/earth-engine/guides/python_install)
  * [Install the gcloud CLI | Google Cloud](https://cloud.google.com/sdk/docs/install)
  * [Python Installation | Google Earth Engine | Google Developers](https://developers.google.com/earth-engine/guides/python_install)
* [Importing Raster Data | Google Earth Engine | Google Developers](https://developers.google.com/earth-engine/guides/image_upload)
```bash
cd /Users/bbest/Library/r-miniconda-arm64/envs/r-reticulate/bin
pip install earthengine-api --upgrade
# in dedicated Terminal
/Users/bbest/Library/r-miniconda-arm64/envs/r-reticulate/bin/earthengine/earthengine authenticate
```

```{python}
import ee
ee.Authenticate()
ee.Initialize()
print(ee.Image("NASA/NASADEM_HGT/001").get("title").getInfo())
```

## Single upload

```{r}
librarian::shelf(
  terra)
  
in_tif  <- "/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/am/am_137098.tif"
out_tif <- "/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/am_137098_a.tif"
tmp_tif <- in_tif
tif     <- "/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/am_137098_b.tif"

datatype = "INT1U"
gdal_datatype <- "Byte"
opt_byte     <- ifelse(datatype == "INT1U", "-a_nodata 255", "")
# OLD pre-COG ----
# opt_compress <- "-co COMPRESS=DEFLATE -co ZLEVEL=9 -co PREDICTOR=2"
# opt_tile     <- "-co TILED=YES" # for COG
# NEW COG ----
# Given an input dataset in.tif with already generated internal or external overviews, a cloud optimized GeoTIFF (COG) can be generated with:
# 
# gdal_translate in.tif out.tif -co TILED=YES -co COPY_SRC_OVERVIEWS=YES -co COMPRESS=LZW
# 
# This will result in a images with tiles of dimension 256x256 pixel for main resolution, and 128x128 tiles for overviews.
opt_compress <- "-co COMPRESS=LZW"
opt_tile     <- "-co TILED=YES -co COPY_SRC_OVERVIEWS=YES" # for COG
opt_sparse   <- "-co SPARSE_OK=TRUE"
opts <- glue::glue("{opt_byte} -ot {gdal_datatype} {opt_compress} {opt_tile} {opt_sparse}")
system(glue::glue("gdal_translate -a_srs EPSG:3857 {opts} '{tmp_tif}' '{tif}'"))

# gdalwarp -t_srs EPSG:3857 -of GTiff -co "COMPRESS=DEFLATE" -co "PREDICTOR=2" $in_tif $out_tif
```

## Upload TIF to GEE Collection

* [Cloud GeoTiff Backed Earth Engine Assets | Google Earth Engine | Google Developers](https://developers.google.com/earth-engine/Earth_Engine_asset_from_cloud_geotiff)
  * [Method: projects.assets.create | Google Earth Engine | Google Developers](https://developers.google.com/earth-engine/reference/rest/v1alpha/projects.assets/create)

## Transfer rasters for GEE upload

Need to explicitly set `EPSG:3857` with updated `write_rast()`.

```{r}
librarian::shelf(
  dplyr, fs, purrr, stringr, terra)
devtools::load_all("~/Github/offshorewindhabitat/offhabr")

dir_tif  <- "/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100"
dir_lyrs <- "/Users/bbest/My Drive/projects/offhab/data/_lyrs"
dir_create(dir_lyrs)

rx_tif <- "([a-z]{2})_([0-9]+)\\.tif$"
d <- tibble(
  path_tif =  dir_ls(dir_tif, regexp=rx_tif, recurse = T)) %>%
  mutate(
    tif       = basename(path_tif),
    ds_key    = str_replace(tif, rx_tif, "\\1"),
    aphia_id  = str_replace(tif, rx_tif, "\\2"),
    lyr_tif   = glue("{dir_lyrs}/{tif}"))
sum(duplicated(basename(d$lyr_tif)))
d$path_tif[1]

d %>% 
  pwalk(
    function(path_tif, lyr_tif, ...){
      write_rast(path_tif, lyr_tif) } )
```

* [TensorFlow for R – gcloud\_install](https://tensorflow.rstudio.com/reference/cloudml/gcloud_install) (source: [gcloud-install.R](https://github.com/rstudio/cloudml/blob/main/R/gcloud-install.R))

```{r}
librarian::shelf(
  cloudml)
 gcloud_install(update = T) 
```


```{python}
import ee
from google.auth.transport.requests import AuthorizedSession

ee.Authenticate()  #  or !earthengine authenticate --auth_mode=gcloud
session = AuthorizedSession(ee.data.get_persistent_credentials())
```

CLOUDSDK_PYTHON=/Users/bbest/Library/r-miniconda-arm64/envs/r-reticulate/bin/python

/Users/bbest/Library/r-miniconda-arm64/envs/r-reticulate/bin:/Users/bbest/google-cloud-sdk/bin:/opt/homebrew/bin:/opt/homebrew/sbin:/usr/local/bin:/System/Cryptexes/App/usr/bin:/usr/bin:/bin:/usr/sbin:/sbin:/Library/Apple/usr/bin:/Users/bbest/Library/r-miniconda-arm64/envs/r-reticulate/bin:/Users/bbest/google-cloud-sdk/bin:/opt/homebrew/bin:/opt/homebrew/sbin:/Users/bbest/Library/r-miniconda-arm64/condabin:/Users/bbest/Applications/quarto/bin:/Library/TeX/texbin:/usr/texbin:/Applications/RStudio.app/Contents/MacOS:/Users/bbest/Library/TinyTeX/bin/universal-darwin


https://samapriya.github.io/geeup/installation/
```bash
pip install geeup
```

https://samapriya.github.io/geeup/projects/image_upload/

https://samapriya.github.io/geeup/projects/cookies_setup/

```bash
geeup cookie_setup
```

```bash
SRC='/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/gee'
MET="$SRC/_metadata.csv"
#DST='ee-offhab/lyrs_ds_aphia'
DST='projects/ee-offhab/assets/lyrs_ds_aphia'
DST='projects/ee-offhab/assets/fldr_ds_aphia'
#geeup getmeta --input $SRC --metadata $MET
geeup upload --source $SRC --dest $DST -m $MET --nodata 255 --pyramids MEAN


# https://developers.google.com/earth-engine/guides/image_manifest
earthengine upload image --manifest ~/gee-test.json "/opt/homebrew/lib/python3.10/site-packages/geeup/data.json"

{"name": "projects/ee-offhab/assets/fldr_ds_aphia/am_341818", "pyramidingPolicy": "MEAN", "properties": {"xsize": 14678, "ysize": 7183, "num_bands": 1, "id_no": "am_341818"}, "missing_data": {"values": [255]}}%      

earthengine rm 'projects/ee-offhab/assets/lyrs_ds_aphia/am_126950'
earthengine upload image --manifest '/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/gee-manifest_am_126950.json'

earthengine rm 'projects/ee-offhab/assets/lyrs_ds_aphia/am_129898'
earthengine upload image --manifest '/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/gee-manifest_am_129898.json'

earthengine rm 'projects/ee-offhab/assets/lyrs_ds_aphia/am_129784'
earthengine upload image --manifest '/Users/bbest/My Drive/projects/offhab/data/_oh_rast0to100/gee-manifest_am_129784.json'
```

## mass upload to gcloud

https://www.tucson.ars.ag.gov/notebooks/uploading_data_2_gee.html

5. Uploading files from local disk into Google Cloud
A single copy can be made from the command line once you are authenticated into Google Cloud.

```bash
sa_json='/Users/bbest/My Drive/private/offhab-google-service-account_09e7228ac965.json'
gcloud auth activate-service-account --key-file $sa_json

cd ~/My\ Drive/projects/offhab/data/_lyrs
# -[m]ultiprocess -[n]ew only
gsutil -m cp -n *.tif gs://offhab_lyrs
```

```{r}
librarian::shelf(
  dplyr, fs, glue, multidplyr, parallel, purrr, stringr)

dir_lyrs <- "/Users/bbest/My Drive/projects/offhab/data/_lyrs"
dir_json <- "/Users/bbest/My Drive/projects/offhab/data/_lyrs_json"
dir_create(dir_json)

rx_tif <- "([a-z]{2})_([0-9]+)\\.tif$"
d <- tibble(
  path_tif =  dir_ls(dir_lyrs, regexp=rx_tif, recurse = T)) %>%
  mutate(
    tif       = basename(path_tif),
    lyr_key   = path_ext_remove(tif),
    ds_key    = str_replace(tif, rx_tif, "\\1"),
    aphia_id  = str_replace(tif, rx_tif, "\\2"),
    path_json   = glue("{dir_json}/{lyr_key}.json")) %>% 
  arrange(lyr_key)

gs2ee <- function(tif, lyr_key, ds_key, aphia_id, path_json, ...){
  # Google Storage (gs) to Earth Engine (ee)
  json <- glue::glue(
    '{{
       "name":"projects/ee-offhab/assets/lyrs_ds_aphia/{lyr_key}",
       "tilesets":[{{"sources":[{{"uris":["gs://offhab_lyrs/{lyr_key}.tif"]}}]}}],
       "pyramidingPolicy":"MEAN",
       "properties":{{
          "lyr_key":"{lyr_key}",
          "ds_key":"{ds_key}",
          "aphia_id":{aphia_id}
       }},
       "missing_data":{{"values":[255]}}
    }}')
  writeLines(json, path_json)
  # earthengine rm 'projects/ee-offhab/assets/lyrs_ds_aphia/am_129898'
  cmd <- glue::glue("earthengine upload image --manifest '{path_json}'")
  r <- try(system(cmd))
  if (inherits(r, "try-error"))
    return(F)
  return(T)
}

# setup cluster to parallelize ----
cl <- new_cluster(detectCores() - 1) # eg my MacBook Air has 8 CPUs, so 7 cores used
cluster_library(
  cl, c(
    "dplyr", "glue", "purrr"))
cluster_assign(
  cl,
  gs2ee  = gs2ee)

d <- d %>%
  partition(cl) %>%
  mutate(
    is_on_ee = pmap_lgl(
      list(tif, lyr_key, ds_key, aphia_id, path_json),
      gs2ee)) %>% 
  collect()

# TODO: check
table(d3$is_on_ee)
table(d3$ds_key)
#   am   oa   rl 
# 9645  641  959
```

