---
title: "geemap"
format: html
---

## Quarto

- [geemap docker](https://hub.docker.com/r/bkavlak/geemap)



```bash
# pull
docker pull bkavlak/geemap:latest

# check
docker image ls

# run
docker run -it --name geemap -p 8888:8888 -p 6006:6006 -v '/Users/bbest/My Drive/projects/offhab/data/_geemap':/geemap/data bkavlak/geemap:latest bash
# WARNING: The requested image's platform (linux/amd64) does not match the detected host platform (linux/arm64/v8) and no specific platform was requested
```

Now inside VM prompt like `root@71129cbd804f:/geemap`:

```bash
jupyter notebook --ip=0.0.0.0 --port=8888 --allow-root
```

To access the notebook, open this file in a browser:
    file:///root/.local/share/jupyter/runtime/nbserver-14-open.html
Or copy and paste one of these URLs:
    http://71129cbd804f:8888/?token=695b7087b9f5be8d5c94d4e78defcdd507a8905e16946f8d
 or http://127.0.0.1:8888/?token=695b7087b9f5be8d5c94d4e78defcdd507a8905e16946f8d


Process
Run the container by typing on bash or PowerShell:

docker run -it --name 'CONTAINER NAME' -p 8888:8888 -p 6006:6006 -v 'VOLUME DIRECTORY':/geemap/data bkavlak/geemap:latest bash
You should change text inside '___' as your preference (Don't include apostrophes ' ')
## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this 

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
