---
title: "Interpolate to BOEM Regions"
output: html_document
date: "2022-09-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DisMap regions

```{r}
librarian::shelf(
  arcpullr, dplyr, ggplot2, raster, mapview)

dm_server <-"https://services2.arcgis.com/C8EMgrsFcRFL6LrL/ArcGIS/rest/services/"
lyr <- "DisMAP_Regions_20220516/FeatureServer/0"
lyr_url <- paste0(dm_server, lyr)
d_lyr <- get_spatial_layer(lyr_url) # table(d_lyr$OARegion)
table(d_lyr$OARegion)

# skipping: Aleutian Islands, Eastern Bering Sea, Gulf of Alaska 
dmregions <- c(
  gmx = "Gulf of Mexico",
  nef = "Northeast US Fall",
  nes = "Northeast US Spring",
  sef = "Southeast US Fall",
  ses = "Southeast US Spring",
  seu = "Southeast US Summer",
  wca = "West Coast Annual 2003-Present", # "West Coast Annual",
  wct = "West Coast Triennial 1977-2004") # West Coast Triennial")
dmregions_rds <- c(
  dmregions[!names(dmregions) %in% c("wca","wct")],
  wca = "West Coast Annual",
  wct = "West Coast Triennial")

ply_dmrgns <- d_lyr %>% 
  group_by(OARegion) %>% 
  filter(OARegion %in% dmregions) %>% 
  summarize(
    geometry = st_union(geoms, is_coverage = T) %>% 
      st_cast("POLYGON")) %>% 
  st_as_sf(crs = 4326) %>% 
  mutate(
    one = 1) %>% 
  st_make_valid() %>% 
  st_simplify(dTolerance = 1) # 1 meter

stopifnot(length(setdiff(dmregions, ply_dmrgns$OARegion))==0)

mapview(ply_dmrgns)

ply_dmrgns_mer <- st_transform(ply_dmrgns, 3857)
r_dmrgns_mer   <- raster(ply_dmrgns_mer, res=2000)

write_sf(ply_dmrgns, "data/ply_dismap_rgns.geojson")
writeRaster(r_dmrgns_mer, "data/ply_dismap_rgns_mer.tif")
# ply_boem_rgns <- read_sf("data/ply_boem_rgns.geojson")
```

## Mapping

Originally ran from `/Users/bbest/Github/ecoquants/OceanAdapt/OceanAdapt.Rproj`, then 
moved script and `data_tif` to `/Users/bbest/My Drive/projects/offhab/data/oceanadapt.rutgers.edu/data_tif`.

https://apps-st.fisheries.noaa.gov/dismap/DisMAP.html

> The predicted biomass density (kg per tow) distribution
based on fishery-independent survey data. The
distribution surface is created by applying the inverse
distance weighting (IDW) interpolation algorithm to the
observations in the survey for each species, regions, and
season combination. The grid size is 2km x 2 km. The IDW
approach smooths over multiple observations to
interpolate the biomass across areas where the survey did
not sample.

```{r}
# libraries ----
librarian::shelf(
  # fasterize, htmltools
  dplyr, fs, glue, here, gstat, leaflet, mapview,
  purrr, raster, readr, sf, stringr, tibble)

select <- dplyr::select
options(readr.show_col_types = F)

# functions ----
basehere <- function(path){
  str_replace(path, paste0(here(), "/"), "")
}

get_sp_yr <- function(sp_key, yr){
  # sp_key="ach-spi1"; yr=2019
  
  r_yr_tif <- glue("{dir_tmp}/{brgn}_{sp_key}_{yr}-1yr.tif")
  
  if (!file.exists(r_yr_tif)){
    
    pts <- pts_dmrgn_mer %>% 
      filter(
        sp_key  == !!sp_key,
        year    == !!yr) 
    
    if (nrow(pts) == 0){
      message(glue("    unavailable: {brgn}_{sp_key} for {yr}"))
      return(NA)
    }
    
    if (nrow(pts %>% filter(wtcpue_cbrt > 0)) == 0){
      message(glue("    all zeros: {brgn}_{sp_key} for {yr}"))
      return(NA)
    }
    
    message(glue("    writing: {basename(r_yr_tif)}"))
    mdl <- gstat(
      formula = wtcpue_cbrt ~ 1, locations = pts, 
      nmin = 10, nmax = 15, maxdist = 200000, set = list(idp = 1))
    r_yr <- interpolate(r_brgn_mer, mdl) %>% 
      mask(ply_brgn_mer) # mapview::mapview(r_yr)
  
    writeRaster(r_yr, r_yr_tif)
  }
  raster(r_yr_tif)
}

# paths ----
dir_tif       <- "~/My Drive/projects/offhab/data/oceanadapt.rutgers.edu/data_tif"
dir_tmp       <- "~/My Drive/projects/offhab/data/oceanadapt.rutgers.edu/tmp"
dir_oa        <- "~/Github/ecoquants/OceanAdapt"
d_spp_csv     <- path(dir_tif, "_spp.csv")
d_spp_rgn_csv <- path(dir_tif, "_spp_rgn.csv")

# species list ----
if (any(!file.exists(d_spp_rgn_csv, d_spp_csv))){
  
  d_spp_rgn_0_csv <- here("data_clean/spplist.csv")
  
  d_spp_rgn <- read_csv(d_spp_rgn_0_csv) %>%
    rename(
      sp_sci = spp,
      sp_cmn = common)
   
  # * get unique species key `sp_key` ----
  sp_sci2key <- function(spp){
    gs <- str_split(spp, "\\W")[[1]][1:2]
    g <- str_sub(gs[1], end=3)
    s <- str_sub(gs[2], end=3)
    str_to_lower(glue("{g}-{s}"))
  }
  
  d_spp <- d_spp_rgn %>%
    group_by(sp_sci) %>%
    summarise(sp_cmn = first(sp_cmn)) %>%
    arrange(sp_sci, sp_cmn) %>%
    mutate(
      sp_key = map_chr(sp_sci, sp_sci2key))
  
  sp_keys_dup <- d_spp$sp_key[duplicated(d_spp$sp_key)]
  
  d_spp <- bind_rows(
    d_spp %>%
      filter(!sp_key %in% sp_keys_dup),
    d_spp %>%
      filter(sp_key %in% sp_keys_dup) %>%
      group_by(sp_key) %>%
      mutate(
        i = row_number(),
        sp_key2 = glue("{sp_key}{i}")) %>%
      select(
        sp_sci, sp_cmn, sp_key = sp_key2) ) %>%
    select(sp_key, sp_sci, sp_cmn) %>%
    arrange(sp_key, sp_sci)
  
  d_spp_rgn <- d_spp_rgn %>%
    left_join(
      d_spp %>%
        select(sp_sci, sp_key),
      by = "sp_sci")
    
  write_csv(d_spp, d_spp_csv)
  write_csv(d_spp_rgn, d_spp_rgn_csv)
}

d_spp     <- read_csv(d_spp_csv)
d_spp_rgn <- read_csv(d_spp_rgn_csv)

# BOEM regions ----

# list.files("data_clean", "dat_exploded.*rds$")

ply_brgns <- read_sf(here("data/ply_boem_rgns.geojson")) # mapview(ply_brgns)
ply_brgns_mer <- st_transform(ply_brgns, 3857)

bregions <- c(
  waor = "Washington/Oregon",       # light green
  scal = "Southern California",     # dark green
  natl = "North Atlantic",          # light blue
  matl = "Mid Atlantic",            # dark blue
  cgmx = "Central Gulf of Mexico")  # pink

brgn_pal <- RColorBrewer::brewer.pal(
  length(bregions), "Paired")
names(brgn_pal) = names(bregions)

brgn2dmrgns <- list(
  cgmx = c("gmx"),
  waor = c("wca"),
  scal = c("wca"),
  natl = c("nef","nes"),
  matl = c("nef","nes","sef","ses","seu"))

# iterate over BOEM regions ----
for (brgn in names(bregions)){ # brgn = names(bregions)[1] # brgn = "cgmx"
  bregion <- bregions[[brgn]]
  message(glue("BOEM REGION: {brgn} ({bregion})"))
  
  # get data from all regions
  stopifnot(brgn %in% names(brgn2dmrgns))
  dmrgns <- brgn2dmrgns[[brgn]]
  d_lst <- lapply(
    dmrgns, function(dmrgn){ # dmrgn = dmrgns[1]
      dmregion_rds <- dmregions_rds[[dmrgn]]
      path_rds <- glue("{dir_oa}/data_clean/dat_exploded{dmregion_rds}.rds")
      stopifnot(file_exists(path_rds))
      d <- readRDS(path_rds) })
  d_dmrgn <- bind_rows(d_lst)

  pts_dmrgn <- d_dmrgn %>% 
    # tibble() %>% 
    mutate(
      year        = as.integer(year),
      wtcpue_cbrt = wtcpue^(1/3)) %>% # take the cube root
    rename(
      sp_sci = spp) %>%
    left_join(
      d_spp %>%
        select(sp_sci, sp_key),
      by = "sp_sci") %>%
    st_as_sf(
      coords = c("lon", "lat"), remove = T, crs = 4326) %>%
    select(
      year, sp_key, wtcpue_cbrt)  %>% 
    filter(!is.na(sp_key)) # gmx nrow: 12,921,639 -> 2,525,473
   # dropping: region, haulid, common, stratum, stratumarea, depth, wtcpue
  pts_dmrgn_mer <- st_transform(pts_dmrgn, 3857)
  
  ply_brgn_mer <- ply_brgns_mer %>%
    filter(RESA_summa == bregion)
  r_brgn_mer <- raster(ply_brgn_mer, res=2000)
  
  # iterate over species ----
  if (!all(dmregions_rds[dmrgns] %in% unique(d_spp_rgn$region))){
    message("WHOAH! why are rgns not found?!")
  }
  sp_keys <- d_spp_rgn %>% 
    filter(
      !flagged,
      sp_key != "na-na",
      region %in% dmregions_rds[dmrgns]) %>% 
    distinct(sp_key) %>% 
    arrange(sp_key) %>% 
    pull(sp_key)
  
  for (sp_key in sp_keys){ # sp_key = sp_keys[1]
    
    d_sp <- d_spp_rgn %>% 
      filter(
        region %in% dmregions_rds[dmrgns], 
        sp_key == !!sp_key)
    message(glue("SPECIES: {sp_key} ({d_sp$sp_sci}) {d_sp$sp_cmn}"), "\n")
    
    yrs <- pts_dmrgn_mer %>% 
      st_drop_geometry() %>% 
      tibble() %>% 
      filter(sp_key == !!sp_key) %>% 
      arrange(year) %>% 
      pull(year) %>% 
      unique()
    if (length(yrs) == 0){
      message("  WHOAH! why zero yrs?! SKIPPING")
      next()
    }
    
    #for (yr in yrs){ # yr = yrs[1]
    for (yr in max(yrs)){ # yr =  max(yrs)
      
      message(glue("  YEAR: {yr}"))
  
      r_yrs_tif <- glue("{dir_tif}/{brgn}_{sp_key}_{yr}.tif")
      
      if (!file.exists(r_yrs_tif)){
        message(glue("    TRYING: {basename(r_yrs_tif)}"))
        
        yrs <- (yr-2):(yr+2)
        wts <- 3 - (abs(yr - yrs)) # 1 2 3 2 1
        lst <- sapply(
          yrs, function(yr){
            get_sp_yr(sp_key, yr) })
        idx <- !is.na(lst)
        if (sum(idx) == 0){
          message("    SKIPPING: no data")
          next()
        }
        message(glue("    WRITING: {basename(r_yrs_tif)}"))
        stk <- stack(lst[idx])
        r_yrs <- raster::weighted.mean(stk, wts[idx], na.rm=F)
        # mapview(r_yrs)
        writeRaster(r_yrs, r_yrs_tif)
      }
    }
  }
}
```

```{r}
# plot(r_idw)
# r_idw <- r_idw^3

# Create a continuous palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = values(r_idw),
  na.color = "transparent", alpha = T)

leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addRasterImage(
    r_idw, colors = pal, project = F) %>%
  addLegend(
    pal = pal,
    values = values(r_idw),
    title = paste0(
      sp_common, 
      "<br>∛wtcpue<br>",
      min(yrs), ":", max(yrs)),
    opacity = 1)
```

## Sum by region

```{r}
librarian::shelf(
  dplyr, fs, glue, here, readr, stringr, tibble, tidyr)
dir_tif <- "~/My Drive/projects/offhab/data/oceanadapt.rutgers.edu/data_tif"

d_spp <- read_csv(path(dir_tif, "_spp.csv"))

# get all tifs
d_tifs <- tibble(
  base_tif = list.files(dir_tif, "^[^_]{1}.*tif$"),
  path_tif = path_norm(path(dir_tif, base_tif))) %>% 
  separate(
    base_tif, 
    c("brgn", "sp_key", "yr", "ext"), sep = "[_\\.]")
# ensure all species found
stopifnot(length(setdiff(d_tifs$sp_key, d_spp$sp_key)) == 0)

# filter by max year
d_tifs_my <- d_tifs %>% 
  group_by(brgn, sp_key) %>% 
  filter(yr == max(yr)) %>% 
  ungroup() 

table(d_tifs_my$yr)
table(d_tifs_my$brgn)
table(d_tifs_my$sp_key)

brgns <- sort(unique(d_tifs_my$brgn))
for (brgn in brgns){ # brgn = brgns[1]
# for (brgn in brgns[4:5]){ # brgn = brgns[1]
  
  bregion <- bregions[[brgn]]
  ply_brgn_mer <- ply_brgns_mer %>%
    filter(RESA_summa == bregion)

  message(brgn)
  
  # stk_grd <- glue("{dir_tif}/_spp-biomass_{brgn}.grd")
  # if (!file.exists(stk_grd)){
  stk <- d_tifs_my %>% 
    filter(brgn == !!brgn) %>% 
    pull(path_tif) %>% 
    stack()
    # writeRaster(stk, stk_grd)
  # }
  # stk <- stack(stk_grd)
  
  message("  biomass")
  r_sum <- sum(stk, na.rm = T)  %>% 
    mask(ply_brgn_mer) # cgmx: 190 layers
  writeRaster(r_sum, glue("{dir_tif}/_sum-biomass_{brgn}.tif"), overwrite=T)
  
  d_stk <- tibble(
    lyr = names(stk) %>% 
      str_replace(fixed("."), "-"),
    avg = cellStats(stk, 'mean'),
    max = cellStats(stk, 'max'),
    sd  = cellStats(stk, 'sd'),
    btm = pmax(avg - sd, 0) ) %>%  # vector of bottom threshold values,
    # n0  = cellStats(stk, function(x, na.rm = T){
    #   sum(x==0, na.rm = na.rm) }) ) %>% 
    separate(
      lyr, remove = F,
      c("brgn", "sp_key", "yr"), sep = "[_\\.]")
  write_csv(d_stk, glue("{dir_tif}/_spp-cellstats_{brgn}.csv"))
  
  message("  nspp")
  stk_pa <- stk >= d_stk$btm # presence-absence (1 or 0) stack of species based on bottom
  r_nspp <- sum(stk_pa, na.rm = T) %>% 
    mask(ply_brgn_mer) # mapview(r_nspp)
  writeRaster(r_nspp, glue("{dir_tif}/_nspp_{brgn}.tif"), overwrite=T)
  
  message("  product: nspp * biomass")
  r_prod <- r_nspp * r_sum # mapview(r_prod)
  writeRaster(r_nspp, glue("{dir_tif}/_prod_{brgn}.tif"), overwrite=T)
}
```

## Species Taxonomy: `worrms`

```{r}
librarian::shelf(
  dplyr, purrr, worrms)

d_spp_my <- d_spp %>% 
  filter(
    sp_key %in% d_tifs_my$sp_key)

# drop weirdos
sp_sci_skip <- c("Fish unident.", "Purple striated anemone", "Unsorted shab", "Crustacea shrimp", "Red striated")
d_spp_my <- d_spp_my %>% 
    filter(
      !sp_sci %in% sp_sci_skip) %>% 
    mutate(
      sp_sci = recode(
        sp_sci, 
        `Tube worm`= "Ficopomatus enigmaticus")) %>% 

# get taxonomic records from World Registry of Marine Species (WorRMS),
#   which only works up to 50 names at a time
d_lst <- list(); i_beg = 1
i_max <- ceiling(nrow(d_spp_my)/50)
for (i in 1:i_max){ # i = 1
  i_beg = (i - 1) * 50 + 1 
  i_end = min(i * 50, nrow(d_spp_my))
  message(glue("wm_records_taxamatch() for {i_beg}:{i_end} of {nrow(d_spp_my)}"))
  
  d_lst[[i]] <- d_spp_my %>% 
    mutate(
      wm_recs = wm_records_taxamatch(sp_sci))
}
d_spp_my_wm <- bind_rows(d_lst) %>% 
  unnest(wm_recs) %>% 
  arrange(sp_key, status)
# get first of duplicates, with preference for status = "accepted"
d_spp_my_wm <- d_spp_my_wm[!duplicated(d_spp_my_wm$sp_key),]

# species without any worms data; oh well
sp_key_na <- setdiff(d_spp_my$sp_key, d_spp_my_wm$sp_key)
d_spp_my %>% 
  filter(sp_key %in% sp_key_na)
#    sp_key  sp_sci                  sp_cmn                 
#    <chr>   <chr>                   <chr>                  
#  1 can-ant Cancer anthonyi         yellow rock crab       
#  2 chr-fus Chrysaora fuscens       NA                     
#  3 cru-shr Crustacea shrimp        crustaceans            
#  4 fis-uni Fish unident.           NA                     
#  5 inv-spp Invertebrata spp.       invertebrate           
#  6 pod-sid Podochela sidneyi       shortfinger neck crab  
#  7 pur-str Purple striated anemone Purple striated anemone
#  8 red-str Red striated            NA                     
#  9 ter-spp Terebratellacea spp.    brachiopod             
# 10 tra-spp Trachypeneus spp.       trachypeneid shrimps   
# 11 uns-sha Unsorted shab           NA                     

write_csv(d_spp_my_wm, path(dir_tif, "_spp_worrms.csv"))

d_spp_attr <- wm_attr_data_(id = d_spp_my_wm$AphiaID)
write_csv(d_spp_attr, path(dir_tif, "_spp_worrms_attr.csv"))

table(d_spp_attr$measurementType)
#            AMBI ecological group                        Body size          Body size (qualitative)                         Brooding 
#                               16                              593                               13                                1 
#                      Development                      Feedingtype                 Functional group    Species importance to society 
#                                1                                5                               36                              650 
# Supporting structure & enclosure 
#                                1 


table(d_spp_my_wm[16])
# kingdom
#  Animalia Chromista 
#       644         1

table(d_spp_my_wm[17])
# phylum
#      Annelida    Arthropoda   Brachiopoda      Chordata      Cnidaria    Ctenophora Echinodermata      Mollusca    Ochrophyta      Porifera 
#             5            84             1           378            39             1            75            57             1             4 
length(table(d_spp_my_wm[17])) 
# 10

table(d_spp_my_wm[18])
# class
#       Actinopteri          Anthozoa        Ascidiacea        Asteroidea Bacillariophyceae          Bivalvia       Cephalopoda         Crinoidea 
#               332                28                 1                45                 1                 8                31                 1 
#        Echinoidea    Elasmobranchii        Gastropoda    Hexactinellida       Holocephali     Holothuroidea          Hydrozoa      Malacostraca 
#                 9                35                17                 3                 1                 8                 2                83 
#       Merostomata            Myxini       Ophiuroidea        Polychaeta         Scyphozoa     Solenogastres         Thaliacea 
#                 1                 4                11                 3                 9                 1                 3 
length(table(d_spp_my_wm[18])) 
# 23

table(d_spp_my_wm[19])
# order ...
length(table(d_spp_my_wm[19])) 
# 101

spp_classes <- d_spp_my_wm$class %>% unique() %>% na.omit() %>% sort()

d_classes <- wm_records_names(spp_classes)

d_classes <- bind_rows(d_classes)

d_classes_cmn <- wm_common_id_(id = d_classes$AphiaID)

d_classes %>% 
  select(AphiaID, class = scientificname) %>% 
  left_join(
    d_classes_cmn %>% 
      filter(language_code == "eng") %>% 
      mutate(
        AphiaID = as.integer(id)) %>% 
      select(AphiaID, vernacular),
    by = "AphiaID") %>% 
  write_csv(path(dir_tif, "_spp_class_vernacular.csv"))

# _spp_class_vernacular.csv -> _spp_class_common.csv: 
#    manually googled and selected

read_csv(path(dir_tif, "_spp_class_common.csv")) %>% 
  # select(-n_taxa) %>% 
  left_join(
    d_spp_my_wm %>% 
      group_by(class) %>% 
      summarize(n_taxa = n()),
    by = "class") %>% 
  write_csv(path(dir_tif, "_spp_class_common.csv"))
```


## Species extinction risk: `rredlist`

* [rredlist • rredlist](https://docs.ropensci.org/rredlist/articles/rredlist.html)
* [API - IUCN Red List of Threatened Species](https://apiv3.iucnredlist.org/spatial)
* [API - IUCN Red List of Threatened Species](https://apiv3.iucnredlist.org/api/v3/docs#regions)

TODO:
- https://github.com/cardosopmb/red/blob/master/R/red.R#L714-L876 \
  Prediction of potential species distributions using maximum entropy (maxent)
  * [red - an R package to facilitate species red list assessments according to the IUCN criteria](https://bdj.pensoft.net/article/20530/)
  * https://github.com/cardosopmb?tab=repositories
  * [‪Pedro Cardoso‬ - ‪Google Scholar‬](https://scholar.google.com/citations?hl=en&user=WyB_hggAAAAJ&view_op=list_works&sortby=pubdate)

```{r}
librarian::shelf(
  rredlist)

rredlist::rl_use_iucn()

Sys.setenv(
  IUCN_REDLIST_KEY = readLines("~/My Drive/private/iucnredlist.org_api_token.txt"))
Sys.getenv("IUCN_REDLIST_KEY")

# alo-sap | Alosa sapidissima | American shad | 158670
# https://apiv3.iucnredlist.org/api/v3/species/Alosa%20sapidissima?token=fb963555580923eb0f0b2778060e5bcbd0cca50c0b21c416ed7ff68da9c25c11
# https://apiv3.iucnredlist.org/api/v3/species/id/191206?token=fb963555580923eb0f0b2778060e5bcbd0cca50c0b21c416ed7ff68da9c25c11

d_rl <- tibble(
  scientificname = d_spp_my_wm$scientificname) %>% 
  mutate(
    rl = map(scientificname, function(x){
      message(x)
      y <- rl_search(name = x)
      y$result }),
    rl_len = map_int(rl, length)) %>% 
  filter(
    d_len > 0) %>% 
  select(-rl_len) %>% 
  unnest(rl)

for (i in 1:nrow(d_spp_my_wm)){ # i = 1
  sp_sci <- d_spp_my_wm$scientificname[i]
  sp_key <- d_spp_my_wm$sp_key[i]
  
  x <- rl_search(name = sp_sci)
  if (length(x$result) > 0){
    message(glue("{i}/{nrow(d_spp_my_wm)} {sp_key} ({sp_sci}): writing csv"))
    write_csv(x$result, glue("{dir_tmp}/rl_{sp_key}.csv"))
  } else {
    message(glue("{i}/{nrow(d_spp_my_wm)} {sp_key} ({sp_sci}): MISSING"))
  }
}

d_rl <- list.files(dir_tmp, "^rl_.*\\.csv$", full.names = T) %>% 
    map_df(~read_csv(.))
write_csv(d_rl, path(dir_tif, "_spp_redlist.csv"))

table(d_rl$category)
table(d_rl$criteria)
```

## `rfishbase`


```{r}
librarian::shelf(
  rfishbase)

fb_tables()

d_spp <- read_csv(glue("{dir_tif}/_spp_worrms.csv"))

View(d_spp)
fb_tbl("species") %>% 
  names()

spp_sci <- d_spp %>% 
  filter(rank == "Species") %>% 
  pull(scientificname)

spp_fb <- fb_tbl("species") %>% 
  mutate(sci_name = paste(Genus, Species)) %>%
  filter(sci_name %in% spp_sci) 
write_csv(spp_fb, glue("{dir_tif}/_spp_fb.csv"))

# http://www.fishbase.us/manual/English/FishbaseThe_FOOD_ITEMS_table.htm
d_est <- rfishbase::estimate(spp_sci)
write_csv(d_est, glue("{dir_tif}/_spp_fb_estimate.csv"))

# http://www.fishbase.org/manual/english/fishbasethe_ecology_table.htm
d_eco <- rfishbase::ecology(spp_sci)
write_csv(d_eco, glue("{dir_tif}/_spp_fb_ecology.csv"))

# http://www.fishbase.org/manual/english/fishbasethe_popgrowth_table.htm
d_popgrowth <- rfishbase::popgrowth(spp_sci)
write_csv(d_popgrowth, glue("{dir_tif}/_spp_fb_popgrowth.csv"))

d_stocks <- rfishbase::stocks(spp_sci)
write_csv(d_stocks, glue("{dir_tif}/_spp_fb_stocks.csv"))
```


## DEM: 3 arc sec via `ncdf4`

* [Coastal Relief Model | NCEI](https://www.ngdc.noaa.gov/mgg/coastal/crm.html)
* [opendap.R](https://rstudio-pubs-static.s3.amazonaws.com/350043_e077809976ee44e19f53f146ad223a60.html)


### test w/ NGDC & GEBCO

```{r}
librarian::shelf(
  dplyr, glue, here, mapview, ncdf4, raster, sf, tibble, tidyr)

d_nc <- "https://www.ngdc.noaa.gov/thredds/dodsC/crm/crm_vol1.nc"
d <- nc_open(d_nc)
names(d$var) # z
names(d$dim) # x, y

ply_brgns <- read_sf(here("data/ply_boem_rgns.geojson")) # mapview(ply_brgns)

ply_brgn <- ply_brgns %>% 
  # filter(assess) %>% pull(rgn) %>% paste(collapse = ", ")
  #   Mid Atlantic, North Atlantic, Central Gulf of Mexico, Southern California, Washington/Oregon
  filter(rgn == "North Atlantic")
b <- st_bbox(ply_brgn)

x <- ncvar_get(d, "x")
y <- ncvar_get(d, "y")
ix <- tibble(
  x = x) %>% 
  rowid_to_column("i") %>% 
  filter(
    x >= b["xmin"],
    x <= b["xmax"]) %>% 
  pull(i)
iy <- tibble(
  y = y) %>% 
  rowid_to_column("i") %>% 
  filter(
    y >= b["ymin"],
    y <= b["ymax"]) %>% 
  pull(i)

z <- ncvar_get(
  d, "z", 
  start = c(min(ix), min(iy)), 
  count = c(length(ix), length(iy)))
r <- raster(
  list(x = x[ix], y = y[iy], z = z),
  crs = 4326) %>% 
  mask(ply_brgn) 
mapview(r) +
  mapview(ply_brgn)
natl_vol1_tif <- glue("~/My Drive/projects/offhab/data/ngdc_crm/natl_vol1.tif")
writeRaster(r, natl_vol1_tif)

## Loading required package: sp
plot(raster(list(x = X[xdx], y = Y[ydx], z = SST)))

ply_brgn <- ply_brgns %>% 
  # filter(assess) %>% pull(rgn) %>% paste(collapse = ", ")
  #   Mid Atlantic, North Atlantic, Central Gulf of Mexico, Southern California, Washington/Oregon
  filter(rgn == "Southern California")
b <- st_bbox(ply_brgn)

gebco_nc = "~/My Drive/projects/bbnj/data/raw/gebco.net_depth/GEBCO_2014_2D.nc"

d <- nc_open(gebco_nc)
names(d$var) # elevation
names(d$dim) # lat, lon
x <- ncvar_get(d, "lon")
y <- ncvar_get(d, "lat")
ix <- tibble(
  x = x) %>% 
  rowid_to_column("i") %>% 
  filter(
    x >= b["xmin"],
    x <= b["xmax"]) %>% 
  pull(i)
iy <- tibble(
  y = y) %>% 
  rowid_to_column("i") %>% 
  filter(
    y >= b["ymin"],
    y <= b["ymax"]) %>% 
  pull(i)

elev <- ncvar_get(
  d, "elevation", 
  start = c(min(ix), min(iy)), 
  count = c(length(ix), length(iy)))
r <- raster(
  list(x = x[ix], y = y[iy], z = elev),
  crs = 4326)
r <- r %>% 
  # mask(r > 0, maskvalue=1) 
  mask(ply_brgn) 
mapview(r) +
  mapview(ply_brgn)
socal_tif <- glue("~/My Drive/projects/offhab/data/gebco.net/socal.tif")
writeRaster(r, socal_tif)
```


### loop brgns over NGDC

```{r}
librarian::shelf(
  dplyr, glue, here, mapview, ncdf4, purrr, raster, 
  rlang, sf, tibble, tidyr)
select = dplyr::select

ply_brgns <- read_sf(here("data/ply_boem_rgns.geojson")) %>% 
  filter(assess) %>% 
  select(
    bregion = `RESA_summa`,
    brgn    = `MMS_PLAN_A`)

# ply_brgn <- ply_brgns %>% 
  # filter(assess) %>% pull(rgn) %>% paste(collapse = ", ")
  #   Mid Atlantic, North Atlantic, Central Gulf of Mexico, Southern California, Washington/Oregon


crm_vol_bbox_sf <- function(vol){
  # vol = 1
  u <- glue("https://www.ngdc.noaa.gov/thredds/dodsC/crm/crm_vol{vol}.nc")
  d <- nc_open(u)
  x <- ncvar_get(d, "x")
  y <- ncvar_get(d, "y")
  
  st_bbox(
    c(xmin = min(x), 
      xmax = max(x), 
      ymin = min(y), 
      ymax = max(y)), 
    crs = st_crs(4326)) %>% 
    st_as_sfc() %>% 
    st_as_sf()
}
# x <- crm_vol_bbox(1)

vols <- 1:8
crm_vols <- map(vols, crm_vol_bbox_sf) %>% 
  bind_rows() %>% 
  select(
    geom = x) %>% 
  mutate(
    vol = vols)
# mapview(crm_vols)

crm_vol_raster <- function(vol, ply){
  b <- st_bbox(ply)
  
  u <- glue("https://www.ngdc.noaa.gov/thredds/dodsC/crm/crm_vol{vol}.nc")
  d <- nc_open(u)
  
  x <- ncvar_get(d, "x")
  y <- ncvar_get(d, "y")
  ix <- tibble(
    x = x) %>% 
    rowid_to_column("i") %>% 
    filter(
      x >= b["xmin"],
      x <= b["xmax"]) %>% 
    pull(i)
  iy <- tibble(
    y = y) %>% 
    rowid_to_column("i") %>% 
    filter(
      y >= b["ymin"],
      y <= b["ymax"]) %>% 
    pull(i)

  if (length(ix) == 0 | length(iy) == 0 )
    return(NA)
  
  z <- ncvar_get(
    d, "z", 
    start = c(min(ix), min(iy)), 
    count = c(length(ix), length(iy)))
  r <- raster(
    list(x = x[ix], y = y[iy], z = z),
    crs = 4326) %>% 
    mask(ply)
}

# TODO: iterate over BOEM regions
for (brgn in ply_brgns$brgn){ # brgn = ply_brgns$brgn[1]

  message(brgn)
  tif <- glue("~/My Drive/projects/offhab/data/ngdc.noaa.gov/crm/{brgn}.tif")
  
  if (file.exists(tif)){
    message("  tif exists, SKIP")
    next()
  }
    
  ply <- ply_brgns %>% 
    filter(brgn == !!brgn)
  
  x_vols <- st_intersects(ply, crm_vols)[[1]]
  
  mapview(ply) +
    mapview(
      crm_vols %>% filter(vol %in% x_vols))
  rs <- list()
  for (i in seq_along(x_vols)){ # vol = x_vols[1]
    vol = x_vols[i]
    message(glue("  crm_vol_raster({vol}, ply)"))
    rs[[i]] <- crm_vol_raster(vol, ply)
  }
  
  message(glue("  mosaic"))
  rs <- rs[!is.na(rs)]
  if (length(rs) > 1){
    r <- exec("mosaic", !!!rs, fun = mean)
  } else {
    r <- rs[[1]]
  }

  message(glue("  write"))
  writeRaster(r, tif)
  
  # mapview(r) +
  #   mapview(ply_brgn)
}
```

## DEM: GEBCO 

```{r}
elev <- ncvar_get(
  d, "elevation", 
  start = c(min(ix), min(iy)), 
  count = c(length(ix), length(iy)))
r <- raster(
  list(x = x[ix], y = y[iy], z = elev),
  crs = 4326)
r <- r %>% 
  # mask(r > 0, maskvalue=1) 
  mask(ply_brgn) 
mapview(r) +
  mapview(ply_brgn)
socal_tif <- glue("~/My Drive/projects/offhab/data/gebco.net/socal.tif")
writeRaster(r, socal_tif)
```




## `sdmpredictors`


## Extract Rasters to Wind Lease Areas

* [exactextractr](https://isciences.gitlab.io/exactextractr/)


